---
title: "E-commerceDB"
author: "Group-8"
date: "2024-02-27"
output: 
  pdf_document:
  extra_dependencies: ["bbm", "threeparttable"]
---

# Load necessary libraries

```{r}

library(knitr)
library(kableExtra)
library(readr)
library(RSQLite)
```

```{r}
con <- dbConnect(RSQLite::SQLite(), "ecommerce.db")

Customers <- read_csv("Files/Customers.csv")
View(Customers)

Products <- read_csv("Files/Products.csv", 
    skip = 1)
View(Products)

Order_details <- read_csv("Files/Order_Details.csv",
    skip = 1)

Reviews <- read_csv("Files/Reviews.csv",
    skip = 1)

Suppliers <- read_csv("Files/Suppliers.csv")
View(Suppliers)

Product_Discounts <- read_csv("Files/Product_Discounts.csv")
View(Product_Discounts)

Product_Category <- read_csv("Files/Product_Category.csv")
View(Product_Category)

Order_Items <- read_csv("Files/Order_Items.csv")
View(Order_Items)
```

```{r}
# To create empty column 


# To paste the foreign key into 


```

```{sql connection=con}

```


# Part 1: Database Design and Implementation

## 1.1 Entity Relationship Diagram

-   Here Insert ERD \*

The E-R diagram above simulates a real-world e-commerce data ecosystem, capturing the detailed relationships between entities and attributes essential for facilitating online transactions. In addition, it provides a comprehensive view of the e-commerce system, which serves as a platform for users to browse products, make purchases, and securely complete their payments.

### 1.1.1.Assumptions

-   The company only distributes products within the United Kingdom (UK).

-   The Currency used is Pound Sterling (GBP).

-   Attributes formats will be aligned with UK standard formats such as date , addresses , names ...etc

### 1.1.2. Entities and Attributes

This section describes and illustrates the entities and their respective attributes of the ER diagram above. \#### 1.1.2.1. Customer Users who previously have at least once purchased products and placed an order. \#### 1.1.2.2. Supplier Vendors who provide products. Represent the source of the product items.

#### 1.1.2.3. Product

Indicates all products in the stock and available for sale.

# Design Considerations

-   Absence of an Order Entity:

-   The model intentionally skips direct order management. Instead, it focuses on product management and customer interactions through reviews and payment methods.Additionally, This consideration will guarantee that products purchased by customers are not tracked or stored by the system to align with privacy policies.

-   Order Entity not considered in this ER design in order to follow best practices by not having to include orderId as part of product table which might affect the overall performance of DB retrieval.

-   Customer Engagement: By including Reviews, the model emphasizes customer engagement and feedback without directly managing transactions.

-   Payment Information: Including Payment_Method without an Order entity suggests a pre-registration of payment preferences or a simplified wallet storage that could be expanded in the future.

## Logical Schema

Customers ($\underline{Customer_id}$, Customer_Email, Cust_F_Name, Cust_L_Name, Phone_Country_Code, Phone_Num, Cust_Street_Name, Cust_Building_Name, Cust_Zip_Code)

Products ($\underline{Product_id}$, *Discount_Code*, *Category_id*, Product_Name, Product_Price, Product_Availability)

Suppliers ($\underline{Supplier_id}$, Supplier_Email, Supplier_Name, Supplier_Status, Sup_Building_Name, Sup_Street_Name, Sup_Zip_Code)

<!-- Order_Details ($\underline{Order_id}$, ${\textit{Customer_id}}$, Order_Date, Order_Total, Order_Status, S_Building_Name, S_Street_Name, S_Zip_Code, Street_Name, B_Building_Name, B_Street_Name, B_Zip_Code, Payment_Type, Payment_Status) -->

<!-- Discounts ($\underline{Discount_Code}$, Discount_Status, Discount_Amount) -->

<!-- Reviews ($\underline{Review_id}$, ${\textit{Product_id}}$, Review_Rating, Review_Timestamp, Review_Text, Review_Likes) -->

<!-- Categories ($\underline{Category_id}$, Category_Name) -->

<!-- Order_Items (${\textit{Order_id}}$, ${\textit{Customer_id}}$, ${\textit{Product_id}}$, Quantity, Unit_Price) -->

<!-- Supplied_Items ($\textit{Supplier_id}$, ${\textit{Product_id}}$, Supply_Contracts, Delivery_Terms, Pricing_Agreements) -->

<!-- Many to Many relation Table of Supplier and Product SupplierProduct ($\underline{Supplier_id,Product_id}$) -->

<!-- Many to Many relation Table of Order_details and Product ProductOrder_details ($\underline{Order_id, Product_id}$) -->
